---
title: "a3_task3_sage_kime"
author: "Sage Kime"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Read in the data & clean names

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```


## Section 1: Size comparisons for red and purple urchins


#### Exploratory data visualization

```{r}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram(color = "purple",
                 fill = "darkorange") +
  facet_wrap(~common_name) +
  labs(x = "Urchin diameter (cm), measured to the nearest 0.5 cm", y = "Frequency of urchin diameter")

```


```{r}
ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Common Name", y = "Urchin diameter (cm)")
```



#### Descriptive statistics table

Create a summary table that contains the following statistics for purple and red urchins (still ignoring site as a variable): mean, median, standard deviation, and sample size.

```{r}
urchins_stats <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean_size = mean(size), median_size = median(size), sd = sd(size), sample_size = n())

urchins_stats %>% 
  kable(col.names = c("Common Name", "Mean size (cm)", "Median size (cm)", "Standard deviation", "Sample size (n)")) %>% 
  kable_styling(full_width = FALSE)
```



**Takeaways:** Overall, red urchins have a greater standard deviation than purple urchins. This means that the size of read urchins has a larger variance and spreads over a larger range of values. I would expect the red urchin distribution curve to be wider than the distribution curve of the purple urchins because of this larger variability in size. It is helpful to note that the red urchin has a mean ~ 2.3 cm larger than the purple urchin - this is a 34% larger mean than the purple urchins which is quite significant. 



## Section 2. Explore purple urchin counts by site and year

#### Purple Urchins subset

In a piped sequence, create stored subset of the urchins data called purple_urchins in which you: 

- Filter to only include observations for purple urchins
- Convert the date to class “Date” 
- Create a new column that only contains the year
- Create a new column called site_full that contains the full name


```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year_new = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "IVEE" ~ "Isla Vista Reef",
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef"
  ))

```




#### Summary table and exploratory graph

```{r}
purple_summary <- purple_urchins %>% 
  count(site_full, year_new)

ggplot(data = purple_summary, aes(x = year_new, y = n)) +
  geom_col(aes(fill = site_full)) +
  facet_wrap(~site_full) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_binned() +
  labs(x = "Year", y = "Total Count of Purple Urchins") +
  theme(legend.position = "none") +
  scale_x_continuous()
```


